\documentclass[../report.tex]{subfiles}

\begin{document}

\section*{Abstract}

Music generation is a well established domain of machine learning, dating at
least as far back as 1989. Since then, many successful implementations have made
use of general sequence modelling techniques such as RNNs, LSTMs, and more
recently, Transformers. Due to the high power and memory requirements of these
models, there is an increasing amount of interest in biological neural networks,
which are intrinsically power-efficient and sparsely connected. Nevertheless,
biological networks exhibit astounding cognitive capabilities, which makes them
particularly relevant in the quest to alleviate the aforementioned issues of
artificial neural networks.

The aim of this project was to investigate sequence modelling capabilities of
spiking neural networks (SNNs), specifically in the context of monophonic
melodies. We find that certain types of SNNs are able to approximate the
accuracy of simple LSTM models, even under several biological constraints. On
the other hand, training a spiking network on commodity hardware remains
relatively impractical, and we explore potential solutions to this problem.
Furthermore, the SNN model suffers from similar issues as an LSTM, such as the
lack of long-term structure in generated melodies.

\section*{Acknowledgements}

First and foremost, I would like to express sincere gratitude to my supervisor,
Prof. Steve Furber, for his continued support over the last 6 months. It has
been a pleasure and an honour to work with someone so knowledgeable and
passionate about their field.

Next, I am thankful to Mr. Adam Perrett and Dr. Andrew Gait from the APT
research group, for guidance during the initial stages of the project. Their
insights were very helpful in making key implementation decisions.

Last but not least, I am endlessly grateful to my parents for enabling me to
study abroad. The past three years at the University of Manchester have been a
life-changing experience, one which would have not been possible without their
support.

\end{document}
